---
title: "Lab 10"
author: "Harshini 19MIA1050"
date: "2023-03-20"
output: html_document
---

## installing packages
```{r}
#install.packages("tm",repos = "http://cran.us.r-project.org")  # for text mining
#install.packages("SnowballC") # for text stemming
#install.packages("wordcloud") # word-cloud generator 
#install.packages("RColorBrewer") # color palettes
#install.packages("syuzhet") # for sentiment analysis
#install.packages("ggplot2") # for plotting graphs
```

### Initializing all libraries
```{r}
library("tm")
library("SnowballC")
library("wordcloud")
library("RColorBrewer")
library("syuzhet")
library("ggplot2")
```

### Read the text file from local machine , choose file interactively
```{r}
text <- read.delim("C:\\Users\\harsh\\OneDrive\\Desktop\\TeamHealthRawDataForDemo.txt", encoding = "utf-8")
```
### Load the data as a corpus
```{r}
TextDoc <- Corpus(VectorSource(text))
```
### Cleaning up the text data

```{r}
#Replacing "/", "@" and "|" with space
toSpace <- content_transformer(function (x , pattern ) gsub(pattern, " ", x))
TextDoc <- tm_map(TextDoc, toSpace, "/")
TextDoc <- tm_map(TextDoc, toSpace, "@")
TextDoc <- tm_map(TextDoc, toSpace, "\\|")
# Convert the text to lower case
TextDoc <- tm_map(TextDoc, content_transformer(tolower))
# Remove numbers
TextDoc <- tm_map(TextDoc, removeNumbers)
# Remove english common stopwords
TextDoc <- tm_map(TextDoc, removeWords, stopwords("english"))
# Remove your own stop word
# specify your custom stopwords as a character vector
TextDoc <- tm_map(TextDoc, removeWords, c("s", "company", "team")) 
# Remove punctuations
TextDoc <- tm_map(TextDoc, removePunctuation)
# Eliminate extra white spaces
TextDoc <- tm_map(TextDoc, stripWhitespace)
# Text stemming - which reduces words to their root form
TextDoc <- tm_map(TextDoc, stemDocument)
```
### Building the term document matrix

```{r}
# Build a term-document matrix
TextDoc_dtm <- TermDocumentMatrix(TextDoc)
dtm_m <- as.matrix(TextDoc_dtm)
# Sort by descearing value of frequency
dtm_v <- sort(rowSums(dtm_m),decreasing=TRUE)
dtm_d <- data.frame(word = names(dtm_v),freq=dtm_v)
# Display the top 5 most frequent words
head(dtm_d, 5)
```

```{r}
# Plot the most frequent words
barplot(dtm_d[1:5,]$freq, las = 2, names.arg = dtm_d[1:5,]$word,
        col ="lightgreen", main ="Top 5 most frequent words",
        ylab = "Word frequencies")
```
### generating the word cloud

```{r}
set.seed(1234)
wordcloud(words = dtm_d$word, freq = dtm_d$freq, min.freq = 5,
          max.words=100, random.order=FALSE, rot.per=0.40, 
          colors=brewer.pal(8, "Dark2"))
```
### Word Association

```{r}
# Find associations 
findAssocs(TextDoc_dtm, terms = c("good","work","health"), corlimit = 0.25)		
```
```{r}
# Find associations for words that occur at least 50 times
findAssocs(TextDoc_dtm, terms = findFreqTerms(TextDoc_dtm, lowfreq = 50), corlimit = 0.25)

```
#### Sentiment Scores

##### Scores for syuzhet factor
```{r}
# regular sentiment score using get_sentiment() function and method of your choice
# please note that different methods may have different scales
syuzhet_vector <- get_sentiment(text, method="syuzhet")
# see the first row of the vector
head(syuzhet_vector)
# see summary statistics of the vector
summary(syuzhet_vector)
```
##### Sentiment scores for bing vector

```{r}

# bing
bing_vector <- get_sentiment(text, method="bing")
head(bing_vector)
summary(bing_vector)
#affin
afinn_vector <- get_sentiment(text, method="afinn")
head(afinn_vector)
summary(afinn_vector)
```
The summary statistics of bing and afinn vectors also show that the Median value of Sentiment scores is above 0 and can be interpreted as the overall average sentiment across the all the responses is positive.

Because these different methods use different scales, it’s better to convert their output to a common scale before comparing them. This basic scale conversion can be done easily using R’s built-in sign function, which converts all positive number to 1, all negative numbers to -1 and all zeros remain 0.

```{r}
#compare the first row of each vector using sign function
rbind(
  sign(head(syuzhet_vector)),
  sign(head(bing_vector)),
  sign(head(afinn_vector))
)
```

# Emotion Classification

To understand this, explore the get_nrc_sentiments function, which returns a data frame with each row representing a sentence from the original file. The data frame has ten columns (one column for each of the eight emotions, one column for positive sentiment valence and one for negative sentiment valence). The data in the columns (anger, anticipation, disgust, fear, joy, sadness, surprise, trust, negative, positive) can be accessed individually or in sets.

```{r}
# run nrc sentiment analysis to return data frame with each row classified as one of the following
# emotions, rather than a score: 
# anger, anticipation, disgust, fear, joy, sadness, surprise, trust 
# It also counts the number of positive and negative emotions found in each row
d<-get_nrc_sentiment(as.character(text))
# head(d,10) - to see top 10 lines of the get_nrc_sentiment dataframe
head (d,10)

```

```{r}
#transpose
td<-data.frame(t(d))
#The function rowSums computes column sums across rows for each level of a grouping variable.
td_new <- data.frame(td)
#Transformation and cleaning
names(td_new)[1] <- "count"
td_new <- cbind("sentiment" = rownames(td_new), td_new)
rownames(td_new) <- NULL
td_new2<-td_new[1:8,]
#Plot One - count of words associated with each sentiment
quickplot(sentiment, data=td_new2, weight=count, geom="bar", fill=sentiment, ylab="count")+ggtitle("Survey sentiments")

```

#### Bar plot showing emotions

```{r}

#Plot two - count of words associated with each sentiment, expressed as a percentage
barplot(
  sort(colSums(prop.table(d[, 1:8]))), 
  horiz = TRUE, 
  cex.names = 0.7, 
  las = 1, 
  main = "Emotions in Text", xlab="Percentage"
)
```

**************** THE END ********************